Memes: an R interface to the MEME Suite

### 2021-03-26 manuscript.Rmd
Grammar edits

`runAme` syntax is identical to `runDreme` in that discriminative designs can be
constructed by calling list entries by name. Data can be visualized using
`ame_plot_heatmap()`, revealing that the E93 motif is strongly enriched at
Decreasing sites, while l(3)neo38, lola, and br are enriched in Increasing
sites, supporting the *de-novo* discovery results (**Figure \@ref(fig:ame-heatmap)**).
Additionally, AME detects several other transcription factor motifs that
distinguish decreasing and increasing sites, providing additional clues to
potential factors that also bind with E93 affect chromatin accessibility.

@@

Using this approach we can deeply examine the properties of the E93 motif within
each chromatin response category. First, by counting the number of E93 motifs
within each category, we demonstrate that Decreasing sites are more likely than
increasing or static sites to contain an E93 motif (**Figure \@ref(fig:fimo)A**). We extend
these observations by rederiving position-weight matrices from sequences
matching the E93 motif within each category, allowing visual inspection of motif
quality across groups (**Figure \@ref(fig:fimo)B**). Notably, differences in quality at base
positions 8-12 appear to distinguish increasing from decreasing motifs, where
decreasing motifs are more likely to have strong A bases at positions 8 and 9,
while E93 motifs from increasing sites are more likely to have a T base pair at
position 12 (**Figure \@ref(fig:fimo)B**). Examination of bulk FIMO scores (where higher
scores represent motifs more similar to the reference) also reveals differences
in E93 motif quality between groups, in particular, E93 motifs from Decreasing
sites have higher scores (**Figure \@ref(fig:fimo)C**). Together, these data demonstrate that
a key distinguishing factor between whether a sites will decrease or increase in
chromatin accessibility following E93 binding is the number and quality of E93
motifs at that site.

@@

# Availability and Future developments:

>>>>

`runAme` syntax is identical to `runDreme` in that discriminative designs can be
constructed by calling list entries by name. Data can be visualized using
`ame_plot_heatmap()`, revealing that the E93 motif is strongly enriched at
Decreasing sites, while l(3)neo38, lola, and br are enriched in Increasing
sites, supporting the *de-novo* discovery results (**Figure \@ref(fig:ame-heatmap)**).
Additionally, AME detects several other transcription factor motifs that
distinguish decreasing and increasing sites, providing additional clues to
potential factors that also bind with E93 to affect chromatin accessibility.

@@

Using this approach we can deeply examine the properties of the E93 motif within
each chromatin response category. First, by counting the number of E93 motifs
within each category, we demonstrate that Decreasing sites are more likely than
increasing or static sites to contain an E93 motif (**Figure \@ref(fig:fimo)A**). We extend
these observations by rederiving position-weight matrices from sequences
matching the E93 motif within each category, allowing visual inspection of motif
quality across groups (**Figure \@ref(fig:fimo)B**). Notably, differences in quality at base
positions 8-12 appear to distinguish increasing from decreasing motifs, where
decreasing motifs are more likely to have strong A bases at positions 8 and 9,
while E93 motifs from increasing sites are more likely to have a T base pair at
position 12 (**Figure \@ref(fig:fimo)B**). Examination of bulk FIMO scores (where higher
scores represent motifs more similar to the reference) also reveals differences
in E93 motif quality between groups, in particular, E93 motifs from Decreasing
sites have higher scores (**Figure \@ref(fig:fimo)C**). Together, these data demonstrate that
a key distinguishing factor between whether a site will decrease or increase in
chromatin accessibility following E93 binding is the number and quality of E93
motifs at that site.

@@

# Availability and Future Developments:
